### 2017-08-21 sections/design.tex
shorten design section

\section*{Design considerations}

Scheduling, group size and venue are important design considerations contributing to the success of a hack week.   
Longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges. 
By spending more time together, participants are more likely to overcome barriers of professional terminology.
On the other hand, events that are too long may lead to fatigue among attendees, resulting in a drop in positive outcomes later in the workshop.
A well-designed hack week will have a clear schedule that limits the number of parallel sessions, in order to avoid decision fatigue, and will balance the duration of taught components and open project work.

A hackweek requires a flexible workspace environment with ample opportunity for re-configuration. 
Participants must have access to rooms that can accommodate lectures combined with interactive exchanges and individual work on laptops. 
Workspaces must also accommodate interactive project work where small teams can gather and work together uninterrupted by other groups.    
Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure. 
However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
Fortunately active learning teaching methods are creating new environments that foster interactivity through modified seating arrangements and integration of computer monitors at group workstations.

Another important design consideration is group size.
If the group is too large, chances for random participant exchanges are reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
If the group is too small, it is unlikely to achieve the desired level of diversity among participants to foster new collaborations across sub-fields and disciplines.
In the past, we have found groups with sizes between 50 and 70 participants to be large enough to encourage a breadth of projects while allowing the workshop to function as a cohesive group.

As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.

Hack week outcomes depend strongly on the interest and engagement of participants.
Some attendees, usually those with a strong background in hackathons and their specific topic of study, arrive with the goal of writing a scientific article.
Other attendees plan to learn a specific topic, such as machine learning, or to analyze a specific data set that relates to the tools covered in the tutorials.
This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible leadership among hack week organizers.