### 2018-06-21 design.tex

\section*{Design considerations}

Several design elements contribute to the success of a hack week (see also SM, Sections 4.1 and 4.2 for practical guidance).
For example, scheduling: longer events allow for a larger taught component, more ambitious projects and cross-disciplinary exchanges.
By spending more time together, participants are more likely to overcome barriers of professional terminology.
But events that are too long may also lead to fatigue, resulting in a drop in positive outcomes later in the workshop.
A well-designed hack week will have a clear schedule limiting the number of parallel sessions, and balancing the duration of taught components and project work.

The space used is also an important consideration. A hack week requires a flexible workspace, that allows re-configuration, accommodating both lectures combined with interactive exchanges and individual work on laptops, as well as project work in small teams (see SM 4.1.3, 4.1.4).
%Universities are an obvious first choice for hosting a hack week given their access to scientific resources, research support and infrastructure.
%However traditional university lecture halls often do not meet the hack week criteria for interactive exchanges.
Fortunately, many universities are experimenting with new types of spaces that allow these kinds of activities, and the adoption of active learning teaching methods \cite{prince2004} has led to the development of modular classrooms, designed for group activities and flexible seating arrangements.

Another design consideration is group size (see SM, 4.1.4).
In a large group, chances for random participant exchanges may be reduced, and knowledge transfer may decline as the workshop fractures into smaller groups, often among participants who already know each other.
If the group is small, participant selection must be especially carefully managed in order to achieve the desired level of diversity among participants to foster new collaborations.
We have found groups with sizes between 40 and 70 participants to work well in enabling a breadth of projects while allowing the workshop to function as a cohesive group, but we encourage organizers to experiment with group size.

%As mentioned above, the balance between pedagogy and open project work depends both on the goals of the workshop and the topics around which the workshop is organized.
%If participants have little shared knowledge, more teaching may be necessary in order to allow participants to effectively communicate with each other.
%In communities where a shared understanding exists, tutorials can be shortened to focus on more advanced or innovative topics, leaving more time for active participation.

Hack week outcomes depend strongly on the interest and engagement of participants.
Some attendees arrive with the goal of writing a scientific article, others plan to learn a specific topic (e.g., machine learning), or to analyze a specific data using the tools covered in the tutorials.
This leads to a wide variety of project types from sandbox-style explorations to focused work efforts.
This breadth of possible outcomes makes it difficult to design for all possible participant goals, and calls for adaptive, flexible design.
The large variety in participant backgrounds and experiences--and the resulting range in personalities and objectives of attendees--requires careful, active facilitation of both taught and project components of a hack week (see SM 4.2.4). A well-advertised and enforced Code of Conduct is a very effective tool for managing expectations about participant interactions (see SM 4.1.7).
Community building is a core component of a hack week, and the design of hack weeks needs to include both very strong personalities and very shy participants. In particular, the presence of impostor syndrome experienced by many participants must be taken into account during workshop design (see also SM, Section 4.2.5 for concrete suggestions).